invoice analysis

### 2020-10-01 analysis/008_subject_exploration.Rmd
co-occurence matrix funder/subject

ggplot(subject_all_years, aes(fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 2))
```

@@

```{r}
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))
funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
# gt
funder_overview_subject %>%
  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author)) %>%
  gt::gt(rowname_col = "top_level") %>%
  gt::cols_label(oa = "OA Articles") %>%
  fmt_percent(
    columns = vars(Author, Agreement, `Fee Waived`, Other),
    decimals = 0) %>%
  cols_width(
    2:6 ~ px(80)) %>%
  cols_align(
    align = "right",
    columns = 2:6
  ) %>%
  cols_width(
    2 ~ px(120)
  ) %>%
    gt::cols_align(
    align = "left",
    columns = 1
  ) %>%
  fmt_number(
    columns = vars(oa),
    decimals = 0
  ) %>%
  data_color(
    columns =3:6,
    colors = scales::col_numeric(
      # custom defined values - notice that order matters!
      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
      domain = NULL
    )
  ) %>%
  tab_spanner(
    label = md("**OA Sponsor Type**"),
    columns = 3:6
  ) %>%
    tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )) %>%
   tab_options(
    column_labels.border.top.color = "white",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3)
      )
```
 correlation funder / oa share

>>>>

ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 2))
```

@@

```{r}
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))
funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
# gt
funder_overview_subject %>%
  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author)) %>%
  gt::gt(rowname_col = "top_level") %>%
  gt::cols_label(oa = "OA Articles") %>%
  fmt_percent(
    columns = vars(Author, Agreement, `Fee Waived`, Other),
    decimals = 0) %>%
  cols_width(
    2:6 ~ px(80)) %>%
  cols_align(
    align = "right",
    columns = 2:6
  ) %>%
  cols_width(
    2 ~ px(120)
  ) %>%
    gt::cols_align(
    align = "left",
    columns = 1
  ) %>%
  fmt_number(
    columns = vars(oa),
    decimals = 0
  ) %>%
  data_color(
    columns =3:6,
    colors = scales::col_numeric(
      # custom defined values - notice that order matters!
      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
      domain = NULL
    )
  ) %>%
  tab_spanner(
    label = md("**OA Sponsor Type**"),
    columns = 3:6
  ) %>%
    tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )) %>%
   tab_options(
    column_labels.border.top.color = "white",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3)
      )
```

## fields per funder

```{r}
field_funder <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
ggplot(field_funder, aes(x = oa_sponsor, 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") +
  theme(strip.text = element_text(face="bold"))
```

 correlation funder / oa share
